---
title: "p8105_hw1_az2356"
author: "Amelia Zhao"
date: "9/19/2019"
output: github_document
---
# This is my markdown file for Homework 1.

# Problem 1

 Creating a data frame comprised of:
 
  - a random sample of size 8 from a standard Normal distribution
  
  - a logical vector indicating whether elements of the sample are greater than 0
  
  - a character vector of length 8
  
  - a factor vector of length 8, with 3 different factor "levels"
  
  
  
Loading the tidyverse lib so we can create a dataframe using tibble:
```{r}
library(tidyverse)
```


Setting the seed so that the random number generator will generate the same numbers each time- keeping things boring and predictable....

``` {r}
set.seed(10)
```

Creating a dataframe as described above:

```{r}
hw1p1_df = tibble(
  numbers = rnorm (8),
  over1 = c(numbers > 0),
  fruitfamily = c("peach", "apple", "lemon", "banana", "koreanmelon", "cantaloupe", "watermelon", "lime"),
  sweetness = factor(c("sweet", "sweet", "sour", "sweetish", "sweetish", "sweet", "sweet", "sour"))
  
)

# Printing the dataframe:
hw1p1_df
```


Taking the mean of each variable in the dataframe:

```{r}
mean(pull(hw1p1_df, numbers))
mean(pull(hw1p1_df, over1))
mean(pull(hw1p1_df, fruitfamily))
mean(pull(hw1p1_df, sweetness))

```

Only the mean for the numeric and logical vectors will run. Complex vectors such as factor and character cannot have their means calculated, they can only be trimmed. 


# Question 2


